---
title: "Exercises – Event History Analysis with R"
subtitle: "Companion to *Event History Analysis with R*, 2nd Edition (Broström)"
author: ""
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    theme: flatly
    embed-resources: true
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    documentclass: article
    geometry: margin=2.5cm
execute:
  echo: true
  eval: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false
#| eval: true
library(eha)
library(survival)
```

---

> **Instructions.** Each chapter in this exercise booklet corresponds to a chapter
> in the textbook. Exercises are labelled **(T)** (Theoretical) or **(R)**
> (R programming). Solutions may be provided separately.
>
> All datasets used below are available in the **eha** package unless otherwise
> noted. Load the package with `library(eha)`.

---

# Event History and Survival Data

*Reference: Chapter 1 of the textbook.*

**1.1 (T)** In your own words, explain the difference between *right censoring*
and *left truncation*. Give a concrete example of each from demographic or
epidemiological research.

**1.2 (T)** A study follows 200 individuals from a given start date. During
the study, 30 individuals withdraw without having experienced the event of
interest. Is this informative or non-informative censoring? Discuss what
the consequences would be if censoring is informative.

**1.3 (T)** Define the concept of *risk set*. Why is the risk set central to
event history analysis?

**1.4 (T)** In the `oldmort` dataset in **eha**, each individual is represented
by one row. Explain the columns `enter`, `exit`, and `event`. What time scale
is used? What does it mean that an individual may have `enter > 60`?

**1.5 (R)** Load the `oldmort` dataset and answer the following:

a. How many individuals are in the study?
b. What is the proportion of censored observations?
c. Compute the mean and median observation time (`exit − enter`) for censored
   and uncensored individuals separately.
d. Create a histogram of observation times and comment on the shape of
   the distribution.

```{r}
#| label: ex1-5
data(oldmort)

# Number of individuals
nrow(oldmort)

# Proportion censored
mean(oldmort$event == 0)

# Observation time
oldmort$time <- oldmort$exit - oldmort$enter
tapply(oldmort$time, oldmort$event, mean)
tapply(oldmort$time, oldmort$event, median)

# Histogram
hist(oldmort$time, xlab = "Observation time (years)",
     main = "Distribution of observation times")
```

**1.6 (R)** Tabulate the number of observations by sex (`sex`) in `oldmort`.
Is the distribution balanced? Compute the censoring proportion separately
for men and women.

**1.7 (T)** A person is born in 1820, enters the study (Sundsvall 1860–1879) at
age 50, and dies at age 72. Explain why this person is *left truncated* and
how it is handled in the `oldmort` structure using columns `enter`, `exit`,
and `event`.

---

# Single Sample Data

*Reference: Chapter 2 of the textbook.*

**2.1 (T)** Let $T$ be a continuous random variable with density function $f(t)$.
Define:

a. The survival function $S(t)$.
b. The hazard function $h(t)$.
c. The cumulative hazard function $H(t)$.

Show that $S(t) = \exp(-H(t))$.

**2.2 (T)** Assume $T \sim \text{Exp}(\lambda)$. Derive $h(t)$ and show that
the exponential distribution has a constant hazard. What is this property
called?

**2.3 (T)** State the Kaplan–Meier estimator and explain why it is a natural
way to estimate $S(t)$ with censored data. Also state Greenwood's formula for
the variance.

**2.4 (T)** Explain what the Nelson–Aalen estimator measures and how it differs
from the Kaplan–Meier estimator. Under what circumstances do they give similar
results?

**2.5 (R)** Using the `oldmort` dataset, estimate the survival function with
the Kaplan–Meier estimator, separately for men and women.

a. Plot both curves in the same graph with 95% confidence bands.
b. Where do the curves differ most?
c. Read off the median survival (median age-at-death) from the curves and
   compare with the actual quartile ages.

```{r}
#| label: ex2-5
library(survival)
data(oldmort)
fit <- survfit(Surv(enter, exit, event) ~ sex, data = oldmort)
plot(fit, col = c("blue", "red"), xlab = "Age", ylab = "Survival",
     conf.int = TRUE)
legend("topright", legend = c("Male", "Female"),
       col = c("blue", "red"), lty = 1)
print(fit)
```

**2.6 (R)** Estimate the Nelson–Aalen estimator for `oldmort` (all individuals).
Compare the cumulative hazard $\hat{H}(t)$ with $-\log \hat{S}_{KM}(t)$.
Are they the same?

```{r}
#| label: ex2-6
fit_km <- survfit(Surv(enter, exit, event) ~ 1, data = oldmort,
                  type = "kaplan-meier")
fit_na <- survfit(Surv(enter, exit, event) ~ 1, data = oldmort,
                  type = "nelson-aalen")
plot(-log(fit_km$surv), fit_na$cumhaz, pch = ".",
     xlab = "-log(KM)", ylab = "Nelson-Aalen")
abline(0, 1, col = "red")
```

**2.7 (T)** The discrete hazard $h_j = P(T = t_j \mid T \geq t_j)$ is used in
nonparametric estimation. Show how the Kaplan–Meier estimator is expressed in
terms of $\hat{h}_j$.

**2.8 (R)** Perform a log-rank test to compare survival between men and women
in `oldmort`. State $H_0$ and $H_1$, and draw a conclusion at the 5%
significance level.

```{r}
#| label: ex2-8
survdiff(Surv(enter, exit, event) ~ sex, data = oldmort)
```

---

# Proportional Hazards and Cox Regression

*Reference: Chapter 3 of the textbook.*

**3.1 (T)** Define the proportional hazards (PH) model. What does the
proportionality assumption imply? Illustrate graphically how hazard curves for
two groups look under the PH assumption.

**3.2 (T)** Write out the Cox partial likelihood for $n$ individuals. Explain
why it is called a *partial* likelihood and what is not estimated.

**3.3 (T)** Explain what the log-rank test and Cox regression have in common.
Under what circumstances do they yield identical conclusions?

**3.4 (T)** Show that if the hazard ratio (HR) is constant over time, this is
equivalent to $\log(-\log S_1(t))$ and $\log(-\log S_2(t))$ being parallel
lines. How can this be used graphically to check the PH assumption?

**3.5 (R)** Fit a Cox regression to `oldmort` with covariate `sex`.

a. What is the estimated hazard ratio (HR) for sex?
b. Construct a 95% confidence interval for HR.
c. Interpret the result demographically or biologically.

```{r}
#| label: ex3-5
fit <- coxreg(Surv(enter, exit, event) ~ sex, data = oldmort)
summary(fit)
```

**3.6 (R)** Check the PH assumption graphically using $\log(-\log(S(t)))$ plots
for sex in `oldmort`.

```{r}
#| label: ex3-6
fit_km <- survfit(Surv(enter, exit, event) ~ sex, data = oldmort)
plot(fit_km, fun = "cloglog", col = c("blue", "red"),
     xlab = "log(time)", ylab = "log(-log(S))")
legend("topleft", legend = c("Male", "Female"),
       col = c("blue", "red"), lty = 1)
```

**3.7 (R)** Compare the results from a log-rank test and Cox regression on
`oldmort` with covariate `sex`. Are the $p$-values similar? What explains any
differences?

**3.8 (T)** Describe how the baseline cumulative hazard function $H_0(t)$ is
estimated (Breslow estimator) after the Cox model coefficients have been
estimated.

---

# Explanatory Variables and Regression

*Reference: Chapter 4 of the textbook.*

**4.1 (T)** Explain the difference between a continuous covariate and a factor
variable in Cox regression. How is a factor variable with $k$ levels coded in
the model?

**4.2 (T)** What is an *interaction* in a regression model? Give an example
from survival analysis where an interaction between sex and age might be
meaningful.

**4.3 (T)** Describe the proportional hazards assumption for a continuous
covariate $x$. What does it mean that the hazard ratio is $e^\beta$ per unit
increase in $x$?

**4.4 (R)** Fit a Cox model to `oldmort` with the covariate `civ` (civil
status). Which reference category is used? How do you interpret the
coefficients?

```{r}
#| label: ex4-4
data(oldmort)
fit <- coxreg(Surv(enter, exit, event) ~ civ, data = oldmort)
summary(fit)
```

**4.5 (R)** Add an interaction term between `sex` and `civ` to the Cox model
on `oldmort`. Is the interaction statistically significant (LR test,
$\alpha = 0.05$)? Interpret the results.

```{r}
#| label: ex4-5
fit0 <- coxreg(Surv(enter, exit, event) ~ sex + civ,  data = oldmort)
fit1 <- coxreg(Surv(enter, exit, event) ~ sex * civ,  data = oldmort)
anova(fit0, fit1)
```

**4.6 (R)** Investigate whether the variable `birthdate` has a linear
relationship with the log-hazard using Martingale residuals (see Section 6.7
of the textbook). Create a plot and comment.

**4.7 (T)** Explain the Akaike Information Criterion (AIC) and how it is
used for model selection in Cox regression. What does AIC penalise for?

**4.8 (R)** Perform a stepwise backward elimination on `oldmort` with
covariates `sex`, `civ`, `ses` (socioeconomic status), and `birthdate`.
Report the final model and its AIC.

```{r}
#| label: ex4-8
fit_full <- coxreg(Surv(enter, exit, event) ~ sex + civ + ses + birthdate,
                   data = oldmort)
drop1(fit_full, test = "Chisq")
```

---

# Poisson Regression

*Reference: Chapter 5 of the textbook.*

**5.1 (T)** State the assumptions of the Poisson distribution and explain why
it is appropriate for modelling counts of deaths in a population over a time
period.

**5.2 (T)** Derive the connection between the exponential distribution's hazard
and the Poisson process: if events in a constant-hazard process are
Poisson-distributed with rate $\lambda$, what distribution does the waiting
time to the first event follow?

**5.3 (T)** In Poisson regression for tabular data, an offset $\log(E_{ij})$
is used, where $E_{ij}$ is exposure time. Why is this necessary, and how
is it interpreted?

**5.4 (R)** Use R's built-in Poisson function to estimate a Poisson regression
model on aggregated data from `oldmort`. Aggregate the number of deaths and
person-years by sex and five-year age group.

```{r}
#| label: ex5-4
data(oldmort)
oldmort$agegroup <- cut(oldmort$enter, breaks = seq(60, 100, by = 5),
                        right = FALSE)
# Aggregate
agg <- aggregate(cbind(event, exit - enter) ~ sex + agegroup,
                 data = oldmort, FUN = sum)
names(agg)[3:4] <- c("deaths", "pyears")
# Poisson model
fit_pois <- glm(deaths ~ sex + agegroup + offset(log(pyears)),
                family = poisson, data = agg)
summary(fit_pois)
```

**5.5 (R)** Compare the coefficients from the Poisson regression in Exercise
5.4 with results from a corresponding Cox regression on individual data.
Are they similar?

**5.6 (T)** Explain the link between the *piecewise constant hazard* (PCH)
model and Poisson regression. How does one choose the number of intervals?

**5.7 (R)** Fit a piecewise constant hazard model to `oldmort` with age
breakpoints at 65, 70, 75, and 80 using `phreg()` in **eha**. Plot the
estimated hazard function.

```{r}
#| label: ex5-7
fit <- phreg(Surv(enter, exit, event) ~ sex,
             dist = "pch", cuts = c(65, 70, 75, 80),
             data = oldmort)
plot(fit)
```

---

# More on Cox Regression

*Reference: Chapter 6 of the textbook.*

**6.1 (T)** Define a *time-varying covariate*. Give one medical and one
demographic example. How does the data structure (one row per person vs.
multiple rows per person) differ from that for a time-fixed covariate?

**6.2 (R)** The `oldmort` dataset does not contain a natural time-varying
covariate, but suppose civil status could change during the study. Describe
(without writing code) how you would restructure the data in
*counting process* format (start–stop format) if such information were
available.

**6.3 (T)** Explain the differences between the Breslow, Efron, and exact
method corrections for *ties* in the Cox partial likelihood. When are the
differences largest?

**6.4 (R)** Fit a Cox model to `oldmort` with `sex` and `ses` and compare
AIC for the Breslow and Efron tie-handling methods.

```{r}
#| label: ex6-4
fit_b <- coxreg(Surv(enter, exit, event) ~ sex + ses,
                method = "breslow", data = oldmort)
fit_e <- coxreg(Surv(enter, exit, event) ~ sex + ses,
                method = "efron",   data = oldmort)
AIC(fit_b)
AIC(fit_e)
```

**6.5 (T)** Define *stratification* in Cox regression. When is it used?
What cannot be estimated in a stratified Cox model?

**6.6 (R)** Fit a stratified Cox model to `oldmort` stratified on `civ`
with covariate `sex`. Compare with an unstratified model. Discuss any
differences in the estimated hazard ratios.

```{r}
#| label: ex6-6
fit_strat    <- coxreg(Surv(enter, exit, event) ~ sex + strata(civ),
                       data = oldmort)
fit_no_strat <- coxreg(Surv(enter, exit, event) ~ sex + civ,
                       data = oldmort)
summary(fit_strat)
summary(fit_no_strat)
```

**6.7 (T)** Explain what Schoenfeld residuals are and how they can be used
to formally test the PH assumption.

**6.8 (R)** Test the PH assumption for `sex` and `ses` in a Cox model on
`oldmort` using `cox.zph()`. Interpret the results.

```{r}
#| label: ex6-8
fit <- coxph(Surv(enter, exit, event) ~ sex + ses, data = oldmort)
zph <- cox.zph(fit)
print(zph)
plot(zph)
```

---

# Register-Based Survival Data

*Reference: Chapter 7 of the textbook.*

**7.1 (T)** Describe the advantages and limitations of register-based data
compared with prospective studies for survival analysis. Name at least two
specific challenges.

**7.2 (T)** In population registers, *person-years* are often used as the
exposure measure. Explain how person-years are computed and why they are a
suitable denominator for calculating mortality rates.

**7.3 (R)** The datasets `swedeaths` and `swepop` in **eha** contain deaths
and population data for Sweden. Compute age- and sex-specific mortality rates.
Visualise them on a log scale.

```{r}
#| label: ex7-3
data(swedeaths)
data(swepop)
library(dplyr)
d <- left_join(swedeaths, swepop, by = c("age", "sex", "year"))
d$rate <- d$deaths / d$pop
library(ggplot2)
ggplot(d, aes(x = age, y = rate, color = sex)) +
  geom_line() +
  scale_y_log10() +
  facet_wrap(~ year) +
  labs(title = "Mortality rates by age and sex",
       y = "Mortality rate (log scale)")
```

**7.4 (T)** What is a *communal covariate*? Give an example and explain why
standard Cox partial likelihood can be incorrect if such a covariate is
treated as individual-specific.

**7.5 (R)** Using **eha**, fit a piecewise exponential (PCH) model with
`tpchreg()` on tabulated population data. Include sex as a covariate and
report the hazard ratios.

**7.6 (T)** Explain the difference between *direct* and *indirect* age
standardisation. In what contexts is each method used?

---

# Parametric Models

*Reference: Chapter 8 of the textbook.*

**8.1 (T)** List at least four parametric distributions used in survival
analysis (e.g. exponential, Weibull, log-normal, Gompertz). For each
distribution, state whether the hazard function is monotonically increasing,
decreasing, constant, or unimodal (hump-shaped).

**8.2 (T)** The Weibull distribution has hazard function
$h(t) = \frac{p}{\lambda}\!\left(\frac{t}{\lambda}\right)^{p-1}$.
What happens to the hazard when $p = 1$? When $p > 1$? When $p < 1$?

**8.3 (T)** Explain the *Accelerated Failure Time* (AFT) model. How does the
interpretation of coefficients in an AFT model differ from that in a PH model?

**8.4 (R)** Fit a Weibull model to `oldmort` with covariate `sex` using
`phreg()` in **eha**. Compare the results (hazard ratio for sex) with Cox
regression.

```{r}
#| label: ex8-4
fit_w   <- phreg(Surv(enter, exit, event) ~ sex, dist = "weibull",
                 data = oldmort)
fit_cox <- coxreg(Surv(enter, exit, event) ~ sex, data = oldmort)
summary(fit_w)
summary(fit_cox)
```

**8.5 (R)** Compare the Weibull model and Cox model on `oldmort` using
AIC. Which model is preferred?

**8.6 (T)** Explain what the Weibull distribution's PH and AFT properties have
in common. Show algebraically that the Weibull model is simultaneously a PH
model and an AFT model.

**8.7 (R)** Fit a log-logistic model (AFT) and a Weibull model (PH) to
`oldmort` with `sex` and `ses` as covariates. Compare AIC. Which assumption
about the shape of the hazard is more realistic for these data?

```{r}
#| label: ex8-7
fit_ll <- aftreg(Surv(enter, exit, event) ~ sex + ses,
                 dist = "loglogistic", data = oldmort)
fit_wb <- phreg(Surv(enter, exit, event)  ~ sex + ses,
                dist = "weibull",     data = oldmort)
AIC(fit_ll)
AIC(fit_wb)
```

**8.8 (R)** Plot the estimated hazard function from the Weibull model in
Exercise 8.4 separately for men and women. Add a Nelson–Aalen estimate
for comparison.

---

# Multivariate Survival Models

*Reference: Chapter 9 of the textbook.*

**9.1 (T)** What is meant by *clustered* survival data? Give one concrete
example from medical research and one from demographic research.

**9.2 (T)** Explain the concept of *frailty* (unobserved heterogeneity).
How is it modelled in a frailty model? What is the most common frailty
distribution?

**9.3 (T)** Why does a standard Cox model (which ignores clustering) underestimate
the variance of the regression coefficients? Describe the *robust sandwich*
variance as an alternative approach.

**9.4 (R)** The `infants` dataset in **eha** contains infant survival data
with families as clusters. Fit a Cox model with `id` clustering (robust
sandwich variance) and compare the standard errors with those from a naive
Cox model.

```{r}
#| label: ex9-4
data(infants)
# Naive model
fit_naive  <- coxph(Surv(enter, exit, event) ~ sex + age + ses,
                    data = infants)
# Robust (clustered) model
fit_robust <- coxph(Surv(enter, exit, event) ~ sex + age + ses +
                      cluster(id), data = infants)
summary(fit_naive)$coefficients
summary(fit_robust)$coefficients
```

**9.5 (R)** Fit a *shared frailty* model to `infants` with Gamma frailty
and `id` as the grouping variable. Estimate the frailty variance and
interpret the result.

```{r}
#| label: ex9-5
fit_frailty <- phreg(Surv(enter, exit, event) ~ sex + age + frailty(id),
                     data = infants, dist = "weibull")
summary(fit_frailty)
```

**9.6 (T)** Explain the difference between the *marginal* and *conditional*
parameterisations in frailty models. What type of hazard ratio does each
method report?

**9.7 (R)** Investigate how `ses` (socioeconomic status) affects infant
mortality in `infants` using a stratified Cox model (stratified on `id`).
Compare with the frailty model.

---

# Causality and Matching

*Reference: Chapter 10 of the textbook.*

**10.1 (T)** Explain the *potential outcomes framework* (Rubin's causal model).
What is the *fundamental problem of causal inference* and why can it not be
solved directly?

**10.2 (T)** What is a *propensity score*? How can it be used in matching to
estimate causal effects? What assumptions are required?

**10.3 (T)** Describe Aalen's *additive hazards model*. How does the
interpretation of coefficients differ from the Cox PH model?

**10.4 (R)** Simulate a simple observational study with 500 individuals where
treatment assignment depends on a covariate $X$ (confounding). Compare a naive
survival comparison between treated and untreated with a propensity-score
matched analysis.

```{r}
#| label: ex10-4
set.seed(42)
n    <- 500
x    <- rnorm(n)
# Treatment depends on x (confounding)
trt  <- rbinom(n, 1, plogis(0.8 * x))
# Survival time depends on x and treatment
time <- rexp(n, rate = exp(0.5 * x - 0.3 * trt))
evnt <- rep(1, n)
df   <- data.frame(time, event = evnt, trt, x)

# Naive comparison
fit_naive <- coxph(Surv(time, event) ~ trt, data = df)
summary(fit_naive)

# Propensity score
ps_model <- glm(trt ~ x, family = binomial, data = df)
df$ps    <- predict(ps_model, type = "response")

# Matching (requires the MatchIt package):
# library(MatchIt)
# m          <- matchit(trt ~ x, method = "nearest", data = df)
# df_matched <- match.data(m)
# fit_matched <- coxph(Surv(time, event) ~ trt, data = df_matched)
# summary(fit_matched)
```

**10.5 (T)** What is *dynamic path analysis* and for what purpose is it used
in event history analysis?

**10.6 (T)** Explain the concept of *collider bias*. Give an example of how
collider bias can be inadvertently introduced in a survival study.

---

# Competing Risks Models

*Reference: Chapter 11 of the textbook.*

**11.1 (T)** What are *competing risks*? Give three concrete examples from
medical or demographic research. Why is it problematic to treat competing risks
as ordinary censoring?

**11.2 (T)** Explain the difference between the *cause-specific hazard* function
and the *subdistribution hazard* function (Fine–Gray). Which is easier to
interpret biologically? Which is most relevant for modelling the *cumulative
incidence function* (CIF)?

**11.3 (T)** Show that the sum of cause-specific hazards gives the total hazard:
$$h(t) = \sum_{j=1}^{J} h_j(t).$$
What does this imply for the interpretation of the Kaplan–Meier estimate
applied to each individual cause?

**11.4 (R)** Estimate cumulative incidence functions (CIF) for two competing
causes using `cuminc()` from the **cmprsk** package. Use the simulated data
below and plot the CIFs.

```{r}
#| label: ex11-4
set.seed(123)
n      <- 300
time   <- rexp(n, 0.1)
# Status: 0 = censored, 1 = cause 1, 2 = cause 2
status <- sample(0:2, n, replace = TRUE, prob = c(0.3, 0.4, 0.3))
grp    <- sample(c("A", "B"), n, replace = TRUE)
library(cmprsk)
fit_cif <- cuminc(ftime = time, fstatus = status, group = grp)
plot(fit_cif, col = 1:4, xlab = "Time", ylab = "Cumulative incidence")
```

**11.5 (T)** Why is $1 - \hat{S}_{KM}(t)$ for a single cause (treating the
other causes as censoring) a *biased* estimate of the cumulative incidence?
Explain intuitively.

**11.6 (R)** Fit a Fine–Gray model for the subdistribution hazard of cause 1
from the simulation in Exercise 11.4, using `crr()` from **cmprsk**. Include
the group variable as a covariate.

```{r}
#| label: ex11-6
x      <- as.numeric(grp == "B")
fit_fg <- crr(ftime = time, fstatus = status, cov1 = x, failcode = 1)
summary(fit_fg)
```

**11.7 (T)** Compare the interpretation of coefficients in a cause-specific Cox
regression with those in a Fine–Gray model. Under what circumstances can they
point in opposite directions?

**11.8 (R)** Investigate whether competing risks arise in `oldmort` by
synthetically splitting `event` into two causes at random. Compare the sum of
the two CIFs with $1 - \hat{S}_{KM}(t)$ for all-cause mortality. What do you
observe?

```{r}
#| label: ex11-8
data(oldmort)
set.seed(99)
oldmort$cause <- ifelse(oldmort$event == 0, 0,
                        sample(1:2, sum(oldmort$event), replace = TRUE)[
                          cumsum(oldmort$event)])
library(cmprsk)
fit_cif2 <- cuminc(ftime  = oldmort$exit - oldmort$enter,
                   fstatus = oldmort$cause)
fit_km2  <- survfit(Surv(exit - enter, event) ~ 1, data = oldmort)
# Compare: sum of CIFs vs 1 - KM
plot(fit_cif2, col = c("blue", "red"),
     xlab = "Time (years)", ylab = "Cumulative incidence")
lines(fit_km2, fun = "event", lty = 2, col = "black")
legend("bottomright",
       legend = c("CIF cause 1", "CIF cause 2", "1 - KM (all-cause)"),
       col = c("blue", "red", "black"), lty = c(1, 1, 2))
```

---

# Appendix: Mathematical Tools

*Reference: Appendix A–B of the textbook.*

**A.1 (T)** Explain what a *likelihood function* is and formulate the maximum
likelihood estimator (MLE). Under what assumptions is the MLE asymptotically
normally distributed?

**A.2 (T)** Formulate the *Wald test*, *likelihood ratio test* (LRT), and
*score test* for testing $H_0\colon \beta = 0$ in a regression model. Are they
always equivalent? Discuss.

**A.3 (T)** What are AIC and BIC? How do they differ in how they penalise model
complexity? In what situations is BIC preferred over AIC?

**A.4 (R)** Simulate $n = 200$ observations from a Weibull distribution with
$\lambda = 1$, $p = 2$ (increasing hazard). Censor 30% of observations at
random. Estimate $\lambda$ and $p$ by MLE via `phreg()`. Do the estimates
agree with the true parameter values? Repeat 500 times and examine the
distribution of the MLEs.

```{r}
#| label: exA-4
library(eha)
set.seed(2024)
simulate_weibull <- function(n = 200, lambda = 1, p = 2,
                             cens_prob = 0.3) {
  t      <- rweibull(n, shape = p, scale = lambda)
  c_time <- runif(n, min = 0,
                  max = quantile(t, 1 - cens_prob))
  time   <- pmin(t, c_time)
  event  <- as.integer(t <= c_time)
  data.frame(time, event)
}
df  <- simulate_weibull()
fit <- phreg(Surv(time, event) ~ 1, dist = "weibull", data = df)
summary(fit)
```

**A.5 (T)** Explain the *identification problem* in the Cox model: why can
the baseline hazard $h_0(t)$ and the coefficient vector $\beta$ not be
estimated separately without additional assumptions in the nonparametric Cox
model?

---

*References*

- Broström, G. (2022). *Event History Analysis with R*, 2nd edition.
  Chapman & Hall/CRC.
- R packages: **eha**, **survival**, **cmprsk**, **MatchIt**.
- Package documentation: `?oldmort`, `?infants`, `?swedeaths` in R.
